---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r include=FALSE}
library(knitr)
library(lubridate)
library(dplyr)
```
The data is loaded
The date variable is transformed into char type for selection.
Also, remove not available data from the set.
```{r}
data<-read.csv("activity.csv")
data$date<-as.character(data$date)
data[is.na(data)]<-0
```

## What is mean total number of steps taken per day?
The total number taken per day is as following graph
```{r}
daily<-group_by(data,date)%>%
  summarize(totalsteps=sum(steps))
hist(daily$totalsteps,xlab="Total Steps per Day")
MeanSteps<-mean(daily$totalsteps)
MedianSteps<-median(daily$totalsteps)
```
The mean value of total number of steps taken per day is `r MeanSteps`, the The median value of total number of steps taken per day is `r MedianSteps`
## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
